---
title: "Project 3: STAT302package Tutorial"
output: rmarkdown::html_vignette
author: Eunice Zhang
vignette: >
  %\VignetteIndexEntry{STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Our package,"STAT302package," includes four functions. The first function is called `my_t.test()`, which performs a sample t-test in R for a two-sided t-test, and one-sided t-test.It can compare the means of one sample to some hypothesized value. `my_t.test` takes four parameters, and return a `list` of four elements, which uses the `lifeExp` data from the object `my_gapminder`, which was exported from the `gapminder` package. The second function is call `my_lm()`, which fits a linear model in R. It also uses the `lifeExp` as the response variable and `gdpPercap` and `continent` as the explanatory variables. This function takes two parameters, and return a `table` with rows for each coefficient (including the `(Intercept)`) and columns for the `Estimate`, `Std. Error`, `t value`, and `Pr(>|t|)`. The third function is `my_knn_cv` using the package `palmerpenguins`, which predict output class `species` using covariates `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g` from `my_penguins`. It outputs a vector of the predicted class $\hat{Y}_{i}$ for all observations and a numeric with the cross-validation misclassification error. The fourth function is `my_rf_cv`. This function predicts `body_mass_g` using covariates `bill_length_mm`, `bill_depth_mm`, and `flipper_length_mm` which are also from the package `palmerpenguins`. It takes one parameter which is `k`,and defines a variable `fold` within the `penguins` data. The vignette demonstrates example usage of all main functions. Please file an issue if you have a request for a tutorial that is not currently included and feel free to reach out to us: ziyil4@uw.edu, zhangm27@uw.edu.You can see the vignette by using the following code:


```{r, eval=FALSE}
# To download the STAT302package, use the code below. 

install.packages("STAT302package")
library(STAT302package)

# Alternatively, you can install the development version directly from GitHub.

#devtools::install_github("bryandmartin/corncob")
#library(STAT302package)

# In the vignette, we demonstrates the usage of the four functions of this package and explains how to use this package. 

# Use this to view the vignette in the corncob HTML help
#help(package = "STAT302package", help_type = "html")

# Use this to view the vignette as an isolated HTML file
#utils::browseVignettes(package = "STAT302package")

```


## T-test
```{r setup}
library(STAT302package)
library(tidyverse)
# extract the data
demonstration_data <- my_gapminder$lifeExp

# p_value cut-off of alpha
alpha = 0.05

# Demonstrate a test of the hypothesis (Two-Sided)
result_ts <- my_t.test(demonstration_data, "two.sides", 60)
result_ts$p_val 

# Demonstrate a test of the hypothesis (less)
result_less <- my_t.test(demonstration_data, "less", 60)
result_less$p_val

# Demonstrate a test of the hypothesis (greater)
result_greater <- my_t.test(demonstration_data, "greater", 60)
result_greater$p_val
```
Interpretation: In the majority of analyses, an alpha of 0.05 is used as the cutoff for significance. If the p-value is less than 0.05, we reject the null hypothesis that there’s no difference between the means and conclude that a significant difference does exist. xxxx

## linear Regression
```{r}
# create the formula used for my_lm
demo_regression <- lm(formula = demonstration_data ~ my_gapminder$gdpPercap + my_gapminder$continent, data = my_gapminder)
demo_regression

# demonstrate the linear regression
output_regression <- my_lm(demo_regression, data = my_gapminder)
output_regression
```
Interpretation: gdpPercap coefficient.

Write the hypothesis test associated with the gdpPercap coefficient.
```{r}
# extract the estimates for gdpPercap
est_gdp <- output_regression[2,4]
est_gdp
```
Interpretation: Carefully interpret the results the gdpPercap hypothesis test using a p-value cut-off of α=0.05.

Plot the Actual vs Fitted
```{r}
# calculate y-hat
beta <- as.matrix(output_regression[,1])
alpha <- model.matrix(demonstration_data ~ my_gapminder$gdpPercap + my_gapminder$continent, data = my_gapminder)
y_hat <- alpha %*% beta

# plot the graph
continent <- my_gapminder$continent
my_df <- data.frame(actual = my_gapminder$lifeExp, fitted = y_hat) 
ggplot(my_df, aes(x = fitted, y = actual, color = continent)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "red", lty = 2) + 
  theme_bw(base_size = 15) +
  labs(x = "Fitted values", y = "Actual values", title = "Actual vs. Fitted") +
  theme(plot.title = element_text(hjust = 0.5))
```
Interpretation: Interpret the Actual vs. Fitted plot and make a statement on what it tells you about model fit.



