---
title: "Project 3: STAT302package Tutorial"
author: Eunice Zhang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--- Begin styling code. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- End styling code. --->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
library(STAT302package)
```

## Introduction 


### T-test
```{r}
# extract the data
demonstration_data <- my_gapminder$lifeExp

# p_value cut-off of alpha
alpha = 0.05

# Demonstrate a test of the hypothesis (Two-Sided)
result_ts <- my_t.test(demonstration_data, "two.sides", 60)
result_ts$p_val 

# Demonstrate a test of the hypothesis (less)
result_less <- my_t.test(demonstration_data, "less", 60)
result_less$p_val

# Demonstrate a test of the hypothesis (greater)
result_greater <- my_t.test(demonstration_data, "greater", 60)
result_greater$p_val
```
Interpretation: In the majority of analyses, an alpha of 0.05 is used as the cutoff for significance. If the p-value is less than 0.05, we reject the null hypothesis that there's no difference between the means and conclude that a significant difference does exist.

### Regression
```{r}

```

